\textcolor{sage}{Does Hospital Leadership Matter? Evidence from the Hospital Readmissions Reduction Program}

### 2024-04-23 Slides/readinggroupMarch2024.tex
Updates from Overleaf

\usepackage{biblatex}

\usepackage{tikz}
\usepackage{verbatim}
\setbeamertemplate{note page}{\pagecolor{yellow!5}\insertnote}
\usetikzlibrary{positioning}
\usetikzlibrary{snakes}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{shapes.misc}
\usetikzlibrary{matrix,shapes,arrows,fit,tikzmark}
\usepackage{amsmath}
\usepackage{mathpazo}
\usepackage{hyperref}
\usepackage{lipsum}
\usepackage{multimedia}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bbm}
\newcolumntype{d}[0]{D{.}{.}{5}}
\usepackage{subfigure}

@@

\setbeamercolor{frametitle}{fg=sage}
\setbeamercolor{title}{fg=sage}
\setbeamertemplate{footline}[frame number]
\setbeamertemplate{navigation symbols}{} 
\setbeamertemplate{itemize items}{$\rightarrow$}
\setbeamercolor{itemize item}{fg=sage}
\setbeamercolor{itemize subitem}{fg=sage}
\setbeamercolor{enumerate item}{fg=sage}
\setbeamercolor{enumerate subitem}{fg=sage}
\setbeamercolor{button}{bg=MyBackground,fg=purple,}

@@

\title[]{\textcolor{sage}{Description of Hospital Leadership Data}}

@@

\begin{frame}{end goal}
    \textcolor{purple}{the purpose of collecting this data is to get a sense of who is in charge of running nonprofit hospitals in the US}

@@

    why?
    \begin{wideitemize}
        \item we don't know a lot about who runs hospitals
        \item we know even less about whether characteristics of these people make a difference 
        \item for this project, I care specifically about the characteristic of clinical experience in executive positions
    \end{wideitemize}
\end{frame}

\section{Tax Form 990}

\subsection{who files?}

\begin{frame}{who files?}
\underline{important acronym:} EIN (Employer Identification Number)

    \centering
    \includegraphics[scale=.7]{Graphics/990_snip_frontpage.PNG}
\end{frame}

\begin{frame}{who files?}
    \centering
    \includegraphics[scale=.7]{Graphics/990_snip_officerspage.PNG}
\end{frame}

\subsection{accessing hospital forms}

\begin{frame}{accessing hospital forms}\label{accessing hospital forms}
    \begin{wideitemize}
        \item completed forms are publicly available on ProPublica's website
        \item through their API, I download a list EINs classified as a hospital
        \begin{itemize}
            \item based on National Taxonomy of Exempt Entities (NTEE) codes: E20, E21, E22
        \end{itemize}
        \item along with EIN, I gather their name, zip, state, and urls to tax form pdfs
        \item years 2009-2015
    \end{wideitemize}

@@

    \begin{block}{data point}
        3,925 EINs \hspace{10pt} \hyperlink{why more?}{\beamergotobutton{why more?}}
    \end{block}
\end{frame}

\subsection{matching EIN to AHA hospitals}

\begin{frame}{matching EIN to AHA hospitals}
    I need to be able to link hospitals to other public data sets that capture hospital information
    \vspace{4mm}
    \begin{wideitemize}
        \item I choose to match to the widely used American Hospital Association survey data
        \item only match-able link I have is hospital name
        \item I restrict matches to being in the same state
    \end{wideitemize}
\end{frame}

\begin{frame}{matching EIN to AHA hospitals}
\centering
   \begin{tikzpicture}[node distance=6mm, every node/.style={align=center, font=\large}]

        % Left box
        \node[draw=diagrampurple, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black] (ahaBox) at (0,0) {%
            \underline{\textbf{\textcolor{black}{AHA Hospitals}}} \\[4pt]
            \begin{tabular}{@{}c@{}}
                {Hosp A} \\
                {Hospital B} \\
                {C Hospital} \\
                {Hosp. of D} \\
                {E} \\
            \end{tabular}
        };

        % Right box
        \node[draw=diagramred, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black, right=of ahaBox] (einBox) {%
            \underline{\textbf{\textcolor{black}{EIN Hospitals}}} \\[4pt]
            \begin{tabular}{@{}c@{}}
                {Hospital B} \\
                {Hospital D} \\
                {H} \\
                {C Hospital} \\
                {Hospital G} \\
            \end{tabular}
        };

    \end{tikzpicture}
\end{frame}

\begin{frame}{matching EIN to AHA hospitals}
\centering
   \begin{tikzpicture}[node distance=6mm, every node/.style={align=center, font=\large}]

        % Left box
        \node[draw=diagrampurple, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black] (ahaBox) at (0,0) {%
            \underline{\textbf{\textcolor{black}{AHA Hospitals}}} \\[4pt]
            \begin{tabular}{@{}c@{}}
                \textcolor{diagrampurple}{Hosp A} \\
                {Hospital B} \\
                {C Hospital} \\
                {Hosp. of D} \\
                \textcolor{diagrampurple}{E} \\
            \end{tabular}
        };

        % Right box
        \node[draw=diagramred, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black, right=of ahaBox] (einBox) {%
            \underline{\textbf{\textcolor{black}{EIN Hospitals}}} \\[4pt]
            \begin{tabular}{@{}c@{}}
                {Hospital B} \\
                {Hospital D} \\
                \textcolor{diagramred}{H} \\
                {C Hospital} \\
                \textcolor{diagramred}{Hospital G} \\
            \end{tabular}
        };

    \end{tikzpicture}
\end{frame}

\begin{frame}{matching EIN to AHA hospitals}
\centering
   \begin{tikzpicture}[node distance=6mm, every node/.style={align=center, font=\large}]

        % Left box
        \node[draw=diagrampurple, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black] (ahaBox) at (0,0) {%
            \underline{\textbf{\textcolor{black}{AHA Hospitals}}} \\[4pt]
            \begin{tabular}{@{}c@{}}
                \textcolor{diagrampurple}{Hosp A} \\
                \textcolor{diagramtan}{Hospital B} \\
                \textcolor{diagramtan}{C Hospital} \\
                {Hosp. of D} \\
                \textcolor{diagrampurple}{E} \\
            \end{tabular}
        };

        % Right box
        \node[draw=diagramred, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black, right=of ahaBox] (einBox) {%
            \underline{\textbf{\textcolor{black}{EIN Hospitals}}} \\[4pt]
            \begin{tabular}{@{}c@{}}
                \textcolor{diagramtan}{Hospital B} \\
                {Hospital D} \\
                \textcolor{diagramred}{H} \\
                \textcolor{diagramtan}{C Hospital} \\
                \textcolor{diagramred}{Hospital G} \\
            \end{tabular}
        };

    \end{tikzpicture}

\vspace{7mm}

    \begin{block}{data point}
        I match 1,041 EINs to AHA hospitals using perfect name matching
    \end{block}
\end{frame}

\begin{frame}{matching EIN to AHA hospitals}
\centering
   \begin{tikzpicture}[node distance=6mm, every node/.style={align=center, font=\large}]

        % Left box
        \node[draw=diagrampurple, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black] (ahaBox) at (0,0) {%
            \underline{\textbf{\textcolor{black}{AHA Hospitals}}} \\[4pt]
            \begin{tabular}{@{}c@{}}
                \textcolor{diagrampurple}{Hosp A} \\
                \textcolor{diagramtan}{Hospital B} \\
                \textcolor{diagramtan}{C Hospital} \\
                \textcolor{diagramteal}{Hosp. of D} \\
                \textcolor{diagrampurple}{E} \\
            \end{tabular}
        };

        % Right box
        \node[draw=diagramred, line width=1.5pt, minimum width=4cm, minimum height=4cm, text=black, right=of ahaBox] (einBox) {%
            \underline{\textbf{\textcolor{black}{EIN Hospitals}}} \\[4pt]
            \begin{tabular}{@{}c@{}}
                \textcolor{diagramtan}{Hospital B} \\
                \textcolor{diagramteal}{Hospital D} \\
                \textcolor{diagramred}{H} \\
                \textcolor{diagramtan}{C Hospital} \\
                \textcolor{diagramred}{Hospital G} \\
            \end{tabular}
        };

    \end{tikzpicture}

    \vspace{7mm}

    \begin{block}{data point}
        manually looking at names, I match 196 more EINs to AHA hospitals
    \end{block}
\end{frame}

\subsection{automated vs. manual collection}

\begin{frame}{automated vs. manual collection}
    I now have a list of EINs I want the pdfs for

    \vspace{4mm}

    \begin{wideitemize}
        \item I automate the download of the pdf URLs for these EINs that I collected from the API
        \item in many cases, the API misses the form even though it exists online
        \item I manually download the forms for which this error exists
    \end{wideitemize}
\end{frame}

\section{Text Extraction and Cleaning}

\subsection{optical character recognition}

\begin{frame}{optical character recognition}
    older pdfs are basically images, which requires using OCR for text recognition

    \vspace{4mm}

    \begin{wideitemize}
        \item the software takes pages and returns a string of text found
    \end{wideitemize}
\end{frame}

\begin{frame}{optical character recognition}
\begin{tikzpicture}
\node (bad) {\includegraphics[scale=.5]{Graphics/990_snip_namesexample.PNG}};
\node (good) [right=of bad] {\includegraphics[scale=.45]{Graphics/r_snip_namesexample.PNG}};

\node (a) [fill=purple!50,draw] at (bad.center) {original pdf};
\node (b) [fill=purple!50,draw] at (good.center) {output from OCR};

@@

\subsection{cleaning raw text output}

\begin{frame}{cleaning raw text output}
algorithmically locate names and associated positions

\vspace{4mm}

\begin{enumerate}
    \item remove numbers and special characters
    \item record full names where at least one (first or last) is found in census most common names data
    \item note when the name is followed by ``MD", ``DO", ``RN", etc. 
    \item locate the position nearest the name found and record it
\end{enumerate}
\end{frame}

\begin{frame}{cleaning raw text output}
    final names data:
    \vspace{4mm}
    \begin{wideitemize}
        \item each row is a name, hospital, year observation
        \item other variables: position, clinical experience
    \end{wideitemize}
\end{frame}

\section{Clinical Experience Characteristic}

\subsection{link names to NPPES}

\begin{frame}{link names to NPPES}
    next goal: verify and get more information on doctors in the data
    \vspace{4mm}
    \begin{wideitemize}
        \item use names to link to NPPES to get NPI number and specialty
        \item some cases are easy: MD in tax data and one name match to NPPES
        \item most are not that easy and require manual sleuthing
    \end{wideitemize}
\end{frame}

\begin{frame}{why more EINs than nonprofit hospitals?}\label{why more?}
    \begin{wideitemize}
        \item various sources quote the number of nonprofit hospitals to be less than 3,000
        \begin{wideitemize}
            \item (add citations)
        \end{wideitemize}
        \item without any limitations to hospital EINs at this point, my 3,925 EINs include hospital associations, fundraisers, and some large clinics that will later be excluded.
    \end{wideitemize}

    \hyperlink{accessing hospital forms}{\beamerreturnbutton{go back}}
\end{frame}

>>>>

\usepackage[backend=biber,style=authoryear,
sorting=ynt,citestyle=authoryear]{biblatex}
\addbibresource{Paper/papercitations.bib}

\usepackage{tikz}
\usepackage{verbatim}
\setbeamertemplate{note page}{\pagecolor{yellow!5}\insertnote}
\usetikzlibrary{positioning}
\usetikzlibrary{snakes}
\usetikzlibrary{calc}
\usetikzlibrary{arrows}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{shapes.misc}
\usetikzlibrary{matrix,shapes,arrows,fit,tikzmark}
\usepackage{amsmath}
\usepackage{mathpazo}
\usepackage{hyperref}
\usepackage{lipsum}
\usepackage{multimedia}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bbm}
\newcolumntype{d}[0]{D{.}{.}{5}}
\usepackage{subfigure}
\usepackage{import}

@@

\setbeamercolor{frametitle}{fg=sage}
\setbeamercolor{title}{fg=sage}
\setbeamertemplate{footline}[frame number]
\setbeamertemplate{navigation symbols}{} 
\setbeamertemplate{itemize items}{$\rightarrow$}
\setbeamercolor{itemize item}{fg=sage}
\setbeamercolor{itemize subitem}{fg=sage}
\setbeamercolor{enumerate item}{fg=sage}
\setbeamercolor{enumerate subitem}{fg=sage}
\setbeamercolor{button}{bg=sage,fg=white}

@@

\title[]{\textcolor{sage}{Does Hospital Leadership Matter? Evidence from the Hospital Readmissions Reduction Program}}
\subtitle[]{A JMP update}

@@

\begin{frame}{Motivation: Not-for-profit Firms}
    What do not-for-profit objective functions look like?
    \begin{itemize}
        \item Maximizing prestige, income, quality/quantity tradeoff?  (\cite{steinberg1986revealed})
    \end{itemize}

    \vspace{3mm}

    Important aspect of this literature: using empirical strategies to analyze how firm behavior reveals attributes of objective function (\cite{sloan2000not})

    \vspace{10mm} \pause

    \begin{block}{This Paper}
    I investigate whether executive leadership team composition reveals anything about nonprofit hospital objective functions/behaviors
    \end{block}
\end{frame}

\begin{frame}{Motivation: Nonprofit Hospitals}
    Relevant 
    \begin{itemize}
        \item 50\% of all US hospitals
        \item Staff 207 beds on average
    \end{itemize}

    \vspace{5mm}

    Unclear what objective function looks like 
    \begin{itemize}
        \item Mixed evidence whether costs, uncompensated care, technology adoption, quality differ from for-profit (\cite{sloan2000not}; \cite{eggleston2008hospital}; \cite{moscelli2018effect}; \cite{moscone2020public})
        \item Maximize some combination of profit and output (\cite{deneffe2002not})
        \item Nonprofit hospitals more likely to adjust patient mix after financial shock (\cite{chang2011nonprofit})
    \end{itemize}
\end{frame}

\begin{frame}{Executive Leadership Teams}
    What hasn't been considered is how characteristics even within nonprofits affect individual hospital objectives
    \begin{wideitemize}
        \item The people actually running the hospital are plausibly important
        \item Characteristic important to hospitals is clinical experience 
    \end{wideitemize}
\end{frame}

\begin{frame}{This Paper}
    \begin{block}{}
        I leverage a large-scale policy affecting hospitals to investigate whether clinical experience on a hospital executive team affects hospital response to potential penalties
    \end{block}
\end{frame}

\begin{frame}{About the Policy}
    Hospital Readmissions Reduction Program: passed 2010, took effect 2012

    \vspace{5mm}

    \begin{wideitemize}
        \item Goal: get hospitals to lower readmissions
        \item Hospitals w/ above average readmission rates during look-back period penalized with lower Medicare reimbursements
        \item Conditions: AMI, pneumonia, heart failure
        \item \textbf{All} hospitals had incentive to lower readmissions
    \end{wideitemize}

    \vspace{5mm}\pause

    I mainly use the timing of the policy as variation, but get data on hospital penalties for heterogeneity analysis from Hospital Compare
\end{frame}

\begin{frame}{About the Clinical Experience Variable}\label{clinexp}
    I scrape publicly available Tax Form 990s to get names and titles of all executives for each hospital from 2009-2015 \hyperlink{execnames}{\beamerbutton{text extraction}}

@@

    Two things I need from this:
    \begin{wideitemize}
        \item Characterize stability of leadership teams over time (limit sample)
        \item Characterize whether any executives are MDs 
        \begin{itemize}
            \item This ends up being a time-invariant hospital categorization due to the above sample limitation
        \end{itemize}
    \end{wideitemize}

    \vspace{5mm}

    sample: 580 hospitals, 13\% with clinical experience
\end{frame}

\begin{frame}{About the Outcomes}
    Readmission rates:
    \begin{itemize}
        \item Targeted outcome of the policy, direct incentives for hospitals to lower 
        \item I create average readmission rate across conditions, weighted by number of patients with each condition
    \end{itemize}

    \vspace{5mm}

    Mortality rates:
    \begin{itemize}
        \item Not the targeted outcome, has been used to investigate whether hospitals are actually improving quality or just adjusting readmissions to avoid penalties
        \item I create average mortality rate across conditions, weighted by number of patients with each condition
    \end{itemize}

@@

    I get hospital level readmission and mortality rates from the Hospital Compare Data
\end{frame}

\begin{frame}{Summary Statistics}\label{sumstatsgraph}
    \begin{figure}
        \centering
        \includegraphics[width=\linewidth]{Objects/weighted_read_mort_graph.pdf}
    \end{figure}

    \hyperlink{sumstats}{\beamerbutton{stats}}
\end{frame}

\begin{frame}{Estimation Strategy}

    Compare hospitals with and without clinical experience before and after penalties began:

    \vspace{3mm}

    \begin{equation*}
    y_{ht} = \alpha_{h} + \delta_t + \beta (clinical_h \times post\_penalty_{ht}) + \epsilon_{ht}
    \end{equation*}

    \vspace{10mm}\pause

    Identifying Assumptions:
    \begin{itemize}
        \item Before the penalty, clinical and non-clinical team hospitals trended similarly along quality dimensions
        \item Previously determined leadership team is exogenous to the financial shock
    \end{itemize}

    \vspace{5mm}

    Threat to identification would have to be a time-varying hospital characteristic correlated with both clinical experience and outcomes
\end{frame}

\begin{frame}{All Hospital Response to Policy}\footnotesize
    \import{Tables}{policy_read_table.tex}
    \import{Tables}{policy_mort_table.tex}
\end{frame}

\begin{frame}{Differential Response in Readmissions}
   \small \import{Tables}{read_results_table.tex}
\end{frame}

\begin{frame}{Differential Response in Mortality}
    \small \import{Tables}{mort_results_table.tex}
\end{frame}

\begin{frame}{Conclusion}
\begin{wideitemize}
    \item While non-clinical teams have higher readmission rates before the penalty, they decrease readmissions at a faster rate after the penalties, though the effect is small
    \item No difference in mortality response
\end{wideitemize}

Now the question is why? 
\begin{itemize}
    \item Are non-clinical teams adjusting patient mix? 
\end{itemize}
\end{frame}

\begin{frame}[noframenumbering]{}\label{execnames}
\hyperlink{clinexp}{\beamerbutton{back}}
\begin{tikzpicture}
\node (bad) {\includegraphics[scale=.5]{Graphics/990_snip_namesexample.PNG}};
\node (good) [right=of bad] {\includegraphics[scale=.45]{Graphics/r_snip_namesexample.PNG}};

\node (a) [fill=purple!50,draw] at (bad.center) {original pdf};
\node (b) [fill=purple!50,draw] at (good.center) {output from text extraction};

@@

\begin{frame}[noframenumbering]{}\label{sumstats}
    \scriptsize \import{Tables}{sumstats_nooutcomes.tex}

    \hyperlink{sumstatsgraph}{\beamerbutton{back}}
\end{frame}
